Understanding models for spatial structure in small-area estimation

### 2023-02-16 src/1_checks/lengthscale-recovery.Rmd
arealutils not bsae, and other updates

fits_one <- lapply(ck_one, FUN = function(x) bsae::ck_stan(x$sf))
lengthscale_summary_one <- lapply(fits_one, FUN = function(fit) summary(fit)$summary["l", ])
df_lengthscale_one <- list_to_df(lengthscale_summary_one)

@@

# ck_three <- sim_ck(grid, nsim = 30, l = 3, save = FALSE)
#
# fits_three <- lapply(ck_three, FUN = function(x) bsae::ck_stan(x$sf))
# lengthscale_summary_three <- lapply(fits_three, FUN = function(fit) summary(fit)$summary["l", ])
# df_lengthscale_three <- list_to_df(lengthscale_summary_three)
#
# ggplot(df_lengthscale_three, aes(x = replicate, y = mean)) +
#   geom_point() +
#   geom_errorbar(aes(ymin = X2.5., ymax = X97.5.)) +
#   geom_hline(yintercept = 3, col = lightblue, size = 1.5)
#
# hist(df_lengthscale_three$mean - 3, breaks = 15)
# mean(df_lengthscale_three$mean - 3)

>>>>

fits_one <- lapply(ck_one, FUN = function(x) arealutils::ck_stan(x$sf))
lengthscale_summary_one <- lapply(fits_one, FUN = function(fit) summary(fit)$summary["l", ])
df_lengthscale_one <- list_to_df(lengthscale_summary_one)

@@

# ck_three <- sim_ck(grid, nsim = 30, l = 3, save = FALSE)
#
# fits_three <- lapply(ck_three, FUN = function(x) arealutils::ck_stan(x$sf))
# lengthscale_summary_three <- lapply(fits_three, FUN = function(fit) summary(fit)$summary["l", ])
# df_lengthscale_three <- list_to_df(lengthscale_summary_three)
#
# ggplot(df_lengthscale_three, aes(x = replicate, y = mean)) +
#   geom_point() +
#   geom_errorbar(aes(ymin = X2.5., ymax = X97.5.)) +
#   geom_hline(yintercept = 3, col = lightblue, size = 1.5)
#
# hist(df_lengthscale_three$mean - 3, breaks = 15)
# mean(df_lengthscale_three$mean - 3)

### 2023-02-16 src/docs_paper/paper.Rmd
arealutils not bsae, and other updates

---
title: Understanding models for spatial structure in small-area estimation
author:
  - name: Adam Howes
    email: ath19@ic.ac.uk
    affiliation: A
    corresponding: ath19@ic.ac.uk
  - name: Harrison Zhu
    affiliation: A
  - name: Yidan Xu
    affiliation: C
  - name: Samir Bhatt
    affiliation: D
  - name: Jeffrey W. Eaton
    affiliation: B
  - name: Seth R. Flaxman
    affiliation: E
address:
  - code: A
    address: Department of Mathematics, Imperial College London
  - code: B
    address: MRC Centre for Global Infectious Disease Analysis, Imperial College London
  - code: C
    address: Department of Statistics, University of Michigan
  - code: D
    address: Department of Public Health, University of Copenhagen
  - code: E
    address: Department of Computer Science, University of Oxford
abstract: |
  |  **Background:** Spatial correlation between small-areas is typically accounted for within a generalised linear mixed model setting using spatial random effects. However, commonly used models for spatial random effects based upon adjacency relations, like the Besag model, make unrealistic and unsatisfying spatial assumptions, which seem particularly dubious for irregular geometries. Furthermore, they do not take into account that across many settings areal data is generated by aggregating continuous point data. Defining spatial random effects which instead correspond to aggregated Gaussian processes gives a more intuitively convincing correlation structure, and allows tools from kernel methods to be used in models for areal data.
  |
  |  **Methods:** In a comprehensive simulation study, we used a proper scoring rule to evaluate the performance of different spatial random effect specifications. We performed these simulations under model misspecification and across a range of vignette and realistic geometries with different regularities. We also compared model performance in estimating district-level HIV prevalence from PHIA household survey data across four countries in sub-Saharan Africa, using cross-validation and spatial cross-validation.
  |
  |  **Results:**
  |
  |  **Conclusions:**

@@

\begin{table}
\small
\begin{tabularx}{\textwidth}{llX}
\toprule
Inferential model & Algorithm & Additional details \\
 \midrule
\hypertarget{I1}{I1}. \textit{Constant} & INLA &
No spatial random effects \label{row:I1} \\
\hypertarget{I2}{I2}. \textit{IID} & INLA &
$\phi \sim \mathcal{N}(0, \tau_\phi^{-1} I_n)$\\
\hypertarget{I3}{I3}. \textit{Besag} & INLA &
$\phi \sim \mathcal{N}(0, (\tau_\phi R^\star)^{-})$ following additional recommendations in \ref{app:freni}  \\
\hypertarget{I4}{I4}. \textit{BYM2} & INLA &
$\phi = \tau_\phi^{-1} \left( \sqrt{1- \pi} \, v + \sqrt{\pi} \, u^\star \right)$ with a $\text{Beta}(0.5, 0.5)$ prior on $\pi \in [0, 1]$ \\
\hypertarget{I5}{I5}. \textit{FCK} & INLA &
$\phi \sim \mathcal{N}(0, \tau_\phi^{-1} K)$ with $K_{ij} = k(c_i, c_j)$ where the length-scale $l$ is fixed \\
\hypertarget{I6}{I6}. \textit{CK} & NUTS & As in model \hypertarget{I5}{I5}, with length-scale prior $l \sim \mathrm{Inv{\text-}Gamma}(a, b)$ \\
\hypertarget{I7}{I7}. \textit{FIK} & INLA & $\phi \sim \mathcal{N}(0, \tau_\phi^{-1} K)$ with $K_{ij} = \frac{1}{10^2} \sum_{l = 1}^{10} \sum_{m = 1}^{10} k \left( s^{(i)}_l, s^{(j)}_m \right)$ \newline with hexagonal integration point spacing and fixed length-scale $l$ \\
\hypertarget{I8}{I8}. \textit{IK} & NUTS & As in model \hypertarget{I7}{I7}, with length-scale prior $l \sim \mathrm{Inv{\text-}Gamma}(a, b)$ \\
\bottomrule
\end{tabularx}
\caption{Each inferential model is implemented as a part of the \textsf{R} package \texttt{bsae} available from \url{github.com/athowes/bsae}.}
\label{tab:inferentialmodels}
\small
\end{table}

@@

Following @stein1999interpolation, for the areal kernel models we used the Matérn kernel $k: \Sc \times \Sc \to \mathbb{R}$ given by
\begin{equation}
k(s, s') =
\frac{1}{2^{\nu - 1}\Gamma(\nu)} \left(\frac{\sqrt{2\nu}\lvert s - s' \rvert}{l}\right)^\nu
B_\nu\left(\frac{\sqrt{2\nu}\lvert s - s' \rvert}{l}\right), \label{eq:matern}
\end{equation}
where $B_\nu$ is the modified Bessel function of the second kind, $|s - s'|$ is the Euclidean distance between $s$ and $s'$, $\nu$ is the smoothness hyperparameter and $l$ is the length-scale hyperparameter on the latitude-longitude scale.
We fixed the smoothness parameter, which is otherwise typically unidentifiable from data, to be $3/2$, matching that used in simulation model \hyperlink{S3}{S3}, and convenient in that it simplifies Equation \ref{eq:matern} to be
\begin{equation}
k(s, s') = \left(1 + \sqrt{3} \lvert s - s' \rvert / l \right) \exp \left(- \sqrt{3} \lvert s - s' \rvert / l \right).
\end{equation}
Taking a similar approach to @best1999bayesian, in models \hyperlink{I5}{I5} and \hyperlink{I7}{I7} we fixed the length-scale such that points an average distance apart in the model have 1\% correlation  a priori, independent of the data.
In models \hyperlink{I6}{I6} and \hyperlink{I8}{I8}, we used a length-scale prior $l \sim \mathrm{Inv{\text-}Gamma}(a, b)$, with parameters $a$ and $b$ chosen for each model such that 1\% of the prior mass is below 0.1 and 1\% of the prior mass is above the maximum distance between points in the model [@betancourt2017robust].
For the integration points we set $L_i = 10$ using a hexagonal spacing structure, as with the central panel of Figure \ref{fig:51} -- though `sf::st_sample` with `type = "hexagonal"` does not guarantee exactly the specified number of samples are returned [@pebesma2018sf].
We are limited to this relatively small number by the computational costs of Model \hyperlink{I8}{I8}.
For Model \hyperlink{I6}{I6} although it is feasible to use a much larger number of integration points, since the Gram matrix is only computed once, we chose not to for comparability.

@@

We assessed each fitted inferential model according to its ability to recover the true underlying value of the probabilities $\rho_i$ at each area in the study region, subject to calibration.
We describe our approach to model assessment in the following sections.
<!-- Analogous approaches may be used to check recovery of the intercept parameter $\beta_0$, precision parameter $\tau_\phi$, and, when relevant, the length-scale $l$. -->

@@

The continuous ranked probability score (CRPS) [@matheson1976scoring] gives an overall measure of forecasting performance, and can be seen as a generalisation of the Brier score [@brier1950verification] to distributional forecasts.
For a given fitted model and area $A_i$ let $\rho_i$ have posterior marginal $f(\rho_i) = p(\rho_i \, | \, y)$ and $\omega_i$ be the true value.
Writing $F$ as the cumulative distribution corresponding to $f$ then
\begin{equation}
\text{CRPS}(f, \omega_i) = \int_{-\infty}^{\infty} (F(\rho_i) - \mathbbm{1} \{\rho_i \geq \omega_i \} )^2 \text{d}\rho_i = \int_{0}^{1} (F(\rho_i) - \mathbbm{1} \{\rho_i \geq \omega_i \} )^2 \text{d}\rho_i, \label{eq:crps}
\end{equation}
where $\mathbbm{1}$ denotes the indicator function and the second equality follows from $0 \leq \rho_i \leq 1$.
The minimum possible value of the CRPS is zero, corresponding to a a point mass correctly forecast at the true value $f(\rho_i) = \delta_{\omega_i}$.
The CRPS is a strictly proper scoring rule [@gneiting2007strictly], ensuring that the uniquely best approach to minimising the score is for the model to report its true probabilistic beliefs about the quantity in question.
The logarithmic score [@good1952rational] given by $\text{LogS}(f, \omega_i) = \log f(\omega_i)$ is another popular strictly proper scoring rule, though in line with the findings of @kruger2020predictive, we found it difficult to compute satisfactorily using kernel density estimation.
\begin{equation}
\text{CRPS}(f, \omega_i) \approx \frac{1}{S} \sum_{s = 1}^S | \rho_i^s - \omega_i | - \frac{1}{2S^2} \sum_{s = 1}^S \sum_{l = 1}^S | \rho_i^s - \rho_i^l |. \label{eq:crpsapprox}
\end{equation}

@@

We assessed the coverage of our estimates using posterior predictive checks based on the quantile $q_i = F(\omega_i)$ of the true value within each marginal posterior predictive distribution, computed using numerical integration.
For calibrated models, over repeated simulations, then $q_i \sim \mathcal{U}[0, 1]$ such that at any given nominal coverage $1 - \alpha$, the proportion of quantile-based credible intervals containing the true value $\omega_i$ is also be $1 - \alpha$.
To check uniformity we used probability integral transform (PIT) histograms [@dawid1984present] and empirical cumulative distribution function (ECDF) difference plots [@aldor2013power; @sailynoja2021graphical].

@@

We used data from the most recent publicly available Population Health Impact Assessment (PHIA) survey in each country.
These surveys utilise a complex design, where each individual $j$ in area $i$ has an unequal probability $\pi_{ij}$ of being included in the sample.
<!-- A two-stage design in which enumeration areas are first drawn from a stratified sample and then households are chosen using equal probability systematic sampling from within each enumeration area, is common. -->
To account for the survey design we used sampling weights $\omega_{ij} = 1/\pi_{ij}$ to adjust the raw data in each district, obtaining the Kish effective sample size $m_i^\star = (\sum_k \omega_{ij})^2 / \sum_k \omega_{ij}^2$ and effective number of cases $y_i^\star$ which may be thought of as what would have been observed had the survey been a simple random sample.

@@

We used a the same models as in Section \ref{sec:simulation} (Table \ref{tab:inferentialmodels}), with one alteration to the likelihood.
Since the effective number of cases and effective sample size may not be integers, we modelled the effective number of cases $y_i^\star \in \mathbb{R}$ according to a generalised binomial distribution $y_i^\star \sim \text{xBin}(m_i^\star, \rho_i)$, where $m_i^\star \in \mathbb{R}$ is the effective sample size and $\rho_i \in [0, 1]$ is the adult (15-49) HIV prevalence.
The working likelihood under this model for $m^\star_i \geq y^\star_i$ is given by
\begin{equation}
p(y_i^\star \, | \, m_i^\star, \rho_i) = 
\frac{\Gamma(m_i^\star + 1)}{\Gamma(y_i^\star + 1) \Gamma(m_i^\star - y_i^\star + 1)}  \rho_i ^{y_i^\star}  (1 - \rho_i)^{(m_i^\star - y_i^\star)}. \label{eq:xbinomial}
\end{equation}

@@

We assessed each model using two cross-validation approaches: (1) a standard leave-one-out cross-validation (LOO-CV), in which at each fold the data from a single district were held-out, and (2) a spatial leave-one-out cross-validation (SLOO-CV), in which at each fold the data from a single district and each of its neighbouring districts were removed.
We performed both types of cross-validation manually, as we found approximations that avoid refitting the model, such as the built-in conditional predictive ordinate statistic in `R-INLA` package [@held2010posterior] and the Pareto smoothed importance sampling [@vehtari2015pareto] approach implemented in the `loo` package [@vehtari2017practical], were inaccurate in this setting.
This might be because individual observations in spatial random effect models can exert high influence toward local parameters.

@@

In the presense of structural dependencies, like spatial dependence, LOO-CV tends to underestimate predictive error [@roberts2017cross].
We counteracted this effect by modifying the spatial leave-one-out (SLOO) approach of @le2014spatial.
During the $i$th fold of SLOO-CV, we leave-out the block $(i, \delta i)$ (Figure \ref{fig:sloo}), increasing the extent to which the training and validation sets are conditionally independent.
From the SLOO posterior predictive distribution $p(y_i \, | \, y_{-(i, \delta i)})$  we analogously computed $\text{SLOO-CRPS}_i$.

@@

AH, HZ were supported by the EPSRC Centre for Doctoral Training in Modern Statistics and Statistical Machine Learning (EP/S023151/1).
AH, JWE were supported by the Bill and Melinda Gates Foundation (OPP1190661).
YX was supported by the Mary Lister McCammon Summer Research Fellowship 2019.
HZ was supported by Cervest Limited.
JWE was supported by UNAIDS and National Institute of Allergy and Infectious Disease of the National Institutes of Health (R01AI136664).
SF was supported by the EPSRC (EP/V002910/1).
This research was supported by the MRC Centre for Global Infectious Disease Analysis (MR/R015600/1), jointly funded by the UK Medical Research Council (MRC) and the UK Foreign, Commonwealth \& Development Office (FCDO), under the MRC/FCDO Concordat program and is also part of the EDCTP2 programme supported by the European Union.

@@

AH thanks Robert Ashton for help with \textsf{R} package development, and Tim Lucas, Elizaveta Semenova, Marta Blangiardo and Oliver Ratmann for providing feedback on earlier drafts of this manuscript.

>>>>

---
title: Understanding models for spatial structure in small-area estimation
author:
  - name: Adam Howes
    email: ath19@ic.ac.uk
    affiliation: A
    corresponding: ath19@ic.ac.uk
  - name: Jeffrey W. Eaton
    affiliation: B
  - name: Seth R. Flaxman
    affiliation: C
address:
  - code: A
    address: Department of Mathematics, Imperial College London
  - code: B
    address: MRC Centre for Global Infectious Disease Analysis, Imperial College London
  - code: C
    address: Department of Computer Science, University of Oxford
abstract: |
  |  **Background:** Spatial correlation between small-areas is typically accounted for within a generalised linear mixed model setting using spatial random effects. However, for some geometries, commonly used models for spatial random effects based upon adjacency relations, like the Besag model, make unrealistic and unsatisfying spatial assumptions. Furthermore, they do not take into account that across many settings areal data is generated by aggregating continuous point data. Defining spatial random effects which instead correspond to aggregated Gaussian processes gives a more intuitively convincing correlation structure, and allows tools from kernel methods to be used in models for areal data.
  |
  |  **Methods:** In a simulation study, we used a proper scoring rule to evaluate the performance of different spatial random effect specifications. We performed these simulations under model misspecification and across a range of vignette and realistic geometries with different regularities. We also compared model performance in estimating district-level HIV prevalence from PHIA household survey data across four countries in sub-Saharan Africa, using cross-validation and spatial cross-validation.
  |
  |  **Results:**
  |
  |  **Conclusions:**

@@

\begin{table}
\small
\begin{tabularx}{\textwidth}{llX}
\toprule
Inferential model & Algorithm & Additional details \\
 \midrule
\hypertarget{I1}{I1}. \textit{Constant} & INLA &
No spatial random effects \label{row:I1} \\
\hypertarget{I2}{I2}. \textit{IID} & INLA &
$\phi \sim \mathcal{N}(0, \tau_\phi^{-1} I_n)$\\
\hypertarget{I3}{I3}. \textit{Besag} & INLA &
$\phi \sim \mathcal{N}(0, (\tau_\phi R^\star)^{-})$ following additional recommendations in \ref{app:freni}  \\
\hypertarget{I4}{I4}. \textit{BYM2} & INLA &
$\phi = \tau_\phi^{-1} \left( \sqrt{1- \pi} \, v + \sqrt{\pi} \, u^\star \right)$ with a $\text{Beta}(0.5, 0.5)$ prior on $\pi \in [0, 1]$ \\
\hypertarget{I5}{I5}. \textit{FCK} & INLA &
$\phi \sim \mathcal{N}(0, \tau_\phi^{-1} K)$ with $K_{ij} = k(c_i, c_j)$ where the length-scale $l$ is fixed \\
\hypertarget{I6}{I6}. \textit{CK} & NUTS & As in model \hypertarget{I5}{I5}, with length-scale prior $l \sim \mathrm{Inv{\text-}Gamma}(a, b)$ \\
\hypertarget{I7}{I7}. \textit{FIK} & INLA & $\phi \sim \mathcal{N}(0, \tau_\phi^{-1} K)$ with $K_{ij} = \frac{1}{10^2} \sum_{l = 1}^{10} \sum_{m = 1}^{10} k \left( s^{(i)}_l, s^{(j)}_m \right)$ \newline with hexagonal integration point spacing and fixed length-scale $l$ \\
\hypertarget{I8}{I8}. \textit{IK} & NUTS & As in model \hypertarget{I7}{I7}, with length-scale prior $l \sim \mathrm{Inv{\text-}Gamma}(a, b)$ \\
\bottomrule
\end{tabularx}
\caption{Each inferential model is implemented as a part of the \textsf{R} package \texttt{arealutils} available from \url{github.com/athowes/arealutils}.}
\label{tab:inferentialmodels}
\small
\end{table}

@@

Following @stein1999interpolation, for the areal kernel models we used the Matérn kernel $k: \Sc \times \Sc \to \mathbb{R}$ given by
\begin{equation}
k(s, s') =
\frac{1}{2^{\nu - 1}\Gamma(\nu)} \left(\frac{\sqrt{2\nu}\lvert s - s' \rvert}{l}\right)^\nu
B_\nu\left(\frac{\sqrt{2\nu}\lvert s - s' \rvert}{l}\right), \label{eq:matern}
\end{equation}
where $B_\nu$ is the modified Bessel function of the second kind, $|s - s'|$ is the Euclidean distance between $s$ and $s'$, $\nu$ is the smoothness hyperparameter and $l$ is the length-scale hyperparameter on the latitude-longitude scale.
We fixed the smoothness parameter, which is otherwise typically unidentifiable from data, to be $3/2$, matching that used in simulation model \hyperlink{S3}{S3}, and convenient in that it simplifies Equation \ref{eq:matern} to be
\begin{equation}
k(s, s') = \left(1 + \sqrt{3} \lvert s - s' \rvert / l \right) \exp \left(- \sqrt{3} \lvert s - s' \rvert / l \right).
\end{equation}
Taking a similar approach to @best1999bayesian, in models \hyperlink{I5}{I5} and \hyperlink{I7}{I7} we fixed the length-scale such that points an average distance apart in the model have 1\% correlation  a priori, independent of the data.
In models \hyperlink{I6}{I6} and \hyperlink{I8}{I8}, we used a length-scale prior $l \sim \mathrm{Inv{\text-}Gamma}(a, b)$, with parameters $a$ and $b$ chosen for each model such that 1\% of the prior mass is below 0.1 and 1\% of the prior mass is above the maximum distance between points in the model [@betancourt2017robust].
For the integration points we set $L_i = 10$ using a hexagonal spacing structure, as with the central panel of Figure \ref{fig:51}.
Though note `sf::st_sample` with `type = "hexagonal"` does not guarantee exactly the specified number of samples are returned [@pebesma2018sf].
We are limited to this relatively small number by the computational costs of Model \hyperlink{I8}{I8}.
For Model \hyperlink{I6}{I6} although it is feasible to use a much larger number of integration points, since the Gram matrix is only computed once, we chose not to for comparability.

@@

We assessed each fitted inferential model according to its ability to recover the true underlying value of the probabilities $\rho_i$ at each area in the study region, as well as the calibration of the model's estimates.
<!-- We describe our approach to model assessment in the following sections. -->
<!-- Analogous approaches may be used to check recovery of the intercept parameter $\beta_0$, precision parameter $\tau_\phi$, and, when relevant, the length-scale $l$. -->

@@

The continuous ranked probability score (CRPS) [@matheson1976scoring] gives an overall measure of forecasting performance, and can be seen as a generalisation of the Brier score [@brier1950verification] to distributional forecasts.
For a given fitted model and area $A_i$ let $\rho_i$ have posterior marginal $f(\rho_i) = p(\rho_i \, | \, y)$ and $\omega_i$ be the true value.
Writing $F$ as the cumulative distribution function corresponding to $f$ then
\begin{equation}
\text{CRPS}(f, \omega_i) = \int_{-\infty}^{\infty} (F(\rho_i) - \mathbbm{1} \{\rho_i \geq \omega_i \} )^2 \text{d}\rho_i = \int_{0}^{1} (F(\rho_i) - \mathbbm{1} \{\rho_i \geq \omega_i \} )^2 \text{d}\rho_i, \label{eq:crps}
\end{equation}
where $\mathbbm{1}$ denotes the indicator function and the second equality follows from $0 \leq \rho_i \leq 1$.
The minimum possible value of the CRPS is zero, corresponding to a a point mass correctly forecast at the true value $f(\rho_i) = \delta_{\omega_i}$.
The CRPS is a strictly proper scoring rule such that the uniquely best approach to minimising the score is for the model to report its true probabilistic beliefs about the quantity in question [@gneiting2007strictly].
The logarithmic score [@good1952rational] given by $\text{LogS}(f, \omega_i) = \log f(\omega_i)$ is another popular strictly proper scoring rule, though in line with the findings of @kruger2020predictive, we found it difficult to compute satisfactorily using kernel density estimation.
In contrast, the CRPS can be evaluated directly using samples $\{\rho_i^s\}_{s = 1}^S \sim p(\rho_i \, | \, y)$ as
\begin{equation}
\text{CRPS}(f, \omega_i) \approx \frac{1}{S} \sum_{s = 1}^S | \rho_i^s - \omega_i | - \frac{1}{2S^2} \sum_{s = 1}^S \sum_{l = 1}^S | \rho_i^s - \rho_i^l |. \label{eq:crpsapprox}
\end{equation}

@@

We assessed the coverage of our estimates using posterior predictive checks based on the quantile $q_i = F(\omega_i)$ of the true value within each marginal posterior predictive distribution, computed using numerical integration.
For calibrated models, over repeated simulations, $q_i \sim \mathcal{U}[0, 1]$ such that at any given nominal coverage $1 - \alpha$, the proportion of quantile-based credible intervals containing the true value $\omega_i$ is also be $1 - \alpha$.
To check uniformity we used probability integral transform (PIT) histograms [@dawid1984present] and empirical cumulative distribution function (ECDF) difference plots [@aldor2013power; @sailynoja2021graphical].

@@

We used data from the most recent publicly available Population Health Impact Assessment (PHIA) survey in each country.
These surveys utilise a complex design, where each individual $j$ in area $i$ has an unequal probability $\pi_{ij}$ of being included in the sample.
A two-stage design in which enumeration areas are first drawn from a stratified sample and then households are chosen using equal probability systematic sampling from within each enumeration area, is common.
To account for the survey design we used sampling weights $w_{ij} = 1 / \pi_{ij}$ to adjust the raw data in each district, obtaining the Kish effective sample size $m_i^\star = (\sum_k w_{ij})^2 / \sum_k w_{ij}^2$ and effective number of cases $y_i^\star$ which may be thought of as what would have been observed had the survey been a simple random sample.

@@

We used the same eight inferential models (Section \ref{sec:simulation}, Table \ref{tab:inferentialmodels}), with a small alteration to the likelihood.
Since the effective number of cases and effective sample size may not be integers, we modelled the effective number of cases $y_i^\star \in \mathbb{R}$ according to a generalised binomial distribution $y_i^\star \sim \text{xBin}(m_i^\star, \rho_i)$, where $m_i^\star \in \mathbb{R}$ is the effective sample size and $\rho_i \in [0, 1]$ is the adult (15-49) HIV prevalence.
The working likelihood under this model for $m^\star_i \geq y^\star_i$ is given by
\begin{equation}
p(y_i^\star \, | \, m_i^\star, \rho_i) = 
\frac{\Gamma(m_i^\star + 1)}{\Gamma(y_i^\star + 1) \Gamma(m_i^\star - y_i^\star + 1)}  \rho_i ^{y_i^\star}  (1 - \rho_i)^{(m_i^\star - y_i^\star)}. \label{eq:xbinomial}
\end{equation}

@@

We assessed each model using two cross-validation approaches: (1) a standard leave-one-out cross-validation (LOO-CV), in which at each fold the data from a single district were held-out, and (2) a spatial leave-one-out cross-validation (SLOO-CV), in which at each fold the data from a single district and each of its neighbouring districts were removed.
<!-- We performed both types of cross-validation manually, as we found approximations that avoid refitting the model, such as the built-in conditional predictive ordinate statistic in `R-INLA` package [@held2010posterior] and the Pareto smoothed importance sampling [@vehtari2015pareto] approach implemented in the `loo` package [@vehtari2017practical], were inaccurate in this setting. -->
<!-- This might be because individual observations in spatial random effect models can exert high influence toward local parameters. -->

@@

In the presense of structural dependencies, like spatial dependence, LOO-CV tends to underestimate predictive error [@roberts2017cross].
To counteract this effect we modified the spatial leave-one-out (SLOO) approach of @le2014spatial by leaving-out the block $(i, \delta i)$ (Figure \ref{fig:sloo}) during the $i$th fold, increasing the extent to which the training and validation sets are conditionally independent.
From the SLOO posterior predictive distribution $p(y_i \, | \, y_{-(i, \delta i)})$  we analogously computed $\text{SLOO-CRPS}_i$.

@@

AH was supported by the EPSRC Centre for Doctoral Training in Modern Statistics and Statistical Machine Learning (EP/S023151/1).
AH, JWE were supported by the Bill and Melinda Gates Foundation (OPP1190661).
JWE was supported by UNAIDS and National Institute of Allergy and Infectious Disease of the National Institutes of Health (R01AI136664).
SF was supported by the EPSRC (EP/V002910/1).
This research was supported by the MRC Centre for Global Infectious Disease Analysis (MR/R015600/1), jointly funded by the UK Medical Research Council (MRC) and the UK Foreign, Commonwealth \& Development Office (FCDO), under the MRC/FCDO Concordat program and is also part of the EDCTP2 programme supported by the European Union.

@@

Yidan Xu undertook the Mary Lister McCammon Summer Research Fellowship 2019 on related work.
AH thanks Robert Ashton for help with \textsf{R} package development, and Harrison Zhu, Theo Rashid, Samir Bhatt, Tim Lucas, Elizaveta Semenova, Marta Blangiardo and Oliver Ratmann for helpful conversations.
